---
title: "tw_man"
author: "Sijia Zhang"
date: "12/2/2021"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## data import
```{r}
library(tidyverse)
library(tidylog)
library(magrittr)
library(reshape2)
syllable <- read_csv("tw_man_syllables_1.csv")
```

### Data filtering

??We now make sure that there are no duplicate measurements within syllables.
```{r}
syllable <- syllable %>%
  group_by(syllable_id) %>%
  #filter(!duplicated(time)) %>%
  mutate(syllable_tone = str_remove_all(syllable_tone, "[NULL|/]")) %>%
  ungroup()
unique(syllable[c("syllable_tone")])
summarise(syllable, syllable_id)
unique(syllable[c("word_id")])
```


## Filtering based on durations

How about syllable durations? Should filter overly long / short.

First adding a syllable duration variable + plotting.

```{r}

syllable %>%
ggplot(aes(x=syllable_duration)) +
  geom_histogram() +
  scale_x_log10() +
  xlim(0,0.6)
```

```{r}
# tails: lower end
syllable %>%
  group_by(syllable_id) %>%
  summarise(syllable_duration=syllable_duration[1]) %>%
  filter(syllable_duration < 0.08) %>%
  ungroup() %>%
ggplot(aes(x=syllable_duration)) +
  geom_histogram() # 0.06 makes sense as lower limit

# tails: upper
syllable %>%
  group_by(syllable_id) %>%
  summarise(syllable_duration=syllable_duration[1]) %>%
  filter(syllable_duration > 0.4) %>%
  ungroup() %>%
ggplot(aes(x=syllable_duration)) +
  geom_histogram() # 0.8 makes sense as upper limit
```

Based on the plot, the lower / upper limits are determined as 0.06 and 0.8 s.

```{r}
syllable <- syllable %>%
  filter(syllable_duration >= 0.06,
         syllable_duration <= 0.8)
```
Remove any syllable with a log duration more than three SDs away from the mean (durations are roughly normally distributed in log space, as shown by the plots above)

```{r}
syllable <- syllable %>% 
  filter(
    log(syllable_duration) > 
      mean(log(syllable_duration)) - 
      3*sd(log(syllable_duration)),
    log(syllable_duration) < 
      mean(log(syllable_duration)) +
      3*sd(log(syllable_duration))
  )
```

plot log duration again
```{r}

syllable %>%
ggplot(aes(x=syllable_duration)) +
  geom_histogram() +
  scale_x_log10()
```
## Filtering based on intensity

How about intensity? Should filter overly high/low.

First adding a intensity variable + plotting.

```{r}

syllable %>%
ggplot(aes(x=syllable_intensity)) +
  geom_histogram() +
  scale_x_log10() +
  xlim(30,100)
```

```{r}
# tails: lower end
syllable %>%
  group_by(syllable_id) %>%
  summarise(syllable_intensity=syllable_intensity[1]) %>%
  filter(syllable_intensity < 45) %>%
  ungroup() %>%
ggplot(aes(x=syllable_intensity)) +
  geom_histogram() # 35 makes sense as lower limit

# tails: upper
syllable %>%
  group_by(syllable_id) %>%
  summarise(syllable_intensity=syllable_intensity[1]) %>%
  filter(syllable_intensity > 84) %>%
  ungroup() %>%
ggplot(aes(x=syllable_intensity)) +
  geom_histogram() # 86 makes sense as upper limit
```

Based on the plot, the lower / upper limits are determined as 35 and 86 dB.

```{r}
syllable <- syllable %>%
  filter(syllable_intensity > 35,
         syllable_intensity < 86)
```
Remove any syllable with a log intensity more than three SDs away from the mean (intensities are roughly normally distributed in log space, as shown by the plots above)

```{r}
syllable <- syllable %>% 
  filter(
    log(syllable_intensity) > 
      mean(log(syllable_intensity)) - 
      3*sd(log(syllable_intensity)),
    log(syllable_intensity) < 
      mean(log(syllable_intensity)) +
      3*sd(log(syllable_intensity))
  )
```

plot log intensity again
```{r}

syllable %>%
ggplot(aes(x=syllable_intensity)) +
  geom_histogram() +
  scale_x_log10() +
  xlim(30,100)
```
Should we filter for duration within speakers? Let's plot duration distributions by speaker.

```{r}
syllable_dur <- syllable %>%
  group_by(speaker, syllable_id) %>%
  summarise(syllable_duration=syllable_duration[1]) %>%
  ungroup()

ggplot(syllable_dur, aes(x=syllable_duration)) +
  facet_wrap(~speaker) +
  geom_histogram(bins=20) +
  scale_x_log10()
```
Should we filter for intensity within speakers? Let's plot intensity distributions by speaker.

```{r}
syllable_int <- syllable %>%
  group_by(speaker, syllable_id) %>%
  summarise(syllable_intensity=syllable_intensity[1]) %>%
  ungroup()

ggplot(syllable_int, aes(x=syllable_intensity)) +
  facet_wrap(~speaker) +
  geom_histogram(bins=20) +
  scale_x_log10()
```

## filter read speech ?
# ```{r}
# syllable <- syllable %>%
#   #rownames_to_column("file") %>% 
#   filter(!str_detect(file, 'r') )
# view(syllable)
# ```

```{r}
unique(syllable[c("syllable_id")])
unique(syllable[c("word_id")])
```


## plotting
First plot...
see if syllable duration and intensity will be different in diff position (word-initial/final)
```{r}
syllable_dur <- syllable %>%
  #rownames_to_column("file") %>% 
  #filter(str_detect(file, 'r') )%>% 
  #filter(!grep('r', file)) %>%
  group_by(position_in_word, syllable_id) %>%
  summarise(syllable_duration=syllable_duration[1]) %>%
  ungroup()


ggplot(syllable_dur, aes(x= factor(position_in_word, level = c('initial', 'final')), y=syllable_duration, fill=position_in_word)) +
  geom_boxplot() +
  ylim(0.06, 0.4) +
  # adding the median separately as a point
  stat_summary(fun=mean, geom="point", shape=16) +
  labs(x="Positions in disyllabic words", y = "Syllable duration")
  #scale_y_log10()+

ggsave("Plot of duration by positsions in words.png")
aggregate(syllable_dur$syllable_duration, list(syllable_dur$position_in_word), FUN=mean)

##reorder final and initial
```

```{r}
syllable_int <- syllable %>%
  group_by(position_in_word, syllable_id) %>%
  summarise(syllable_intensity=syllable_intensity[1]) %>%
  ungroup()

ggplot(syllable_int, aes(x=factor(position_in_word, level = c('initial', 'final')), y=syllable_intensity, fill=position_in_word)) +
  geom_boxplot() +
  #ylim(0.06, 0.4) +
  stat_summary(fun=mean, geom="point", shape=16) +
  labs(x="Positions in disyllabic words", y = "Intensity")
  #scale_y_log10()+

ggsave("Plot of intensity by positions in words.png")

aggregate(syllable_int$syllable_intensity, list(syllable_int$position_in_word), FUN=mean)
```
So as expected, duration and intensity does not have a significant difference depending on the syllable position in Taiwannese Mandarin

Now plot syllable duration and intensity by different tone values
Duration:
```{r}
dur_by_tone <- syllable %>%
  group_by(syllable_tone, syllable_id) %>%
  summarise(syllable_duration=syllable_duration[1]) %>%
  ungroup()

ggplot(dur_by_tone, aes(x=syllable_tone, y=syllable_duration, fill=syllable_tone)) +
  geom_boxplot() +
  ylim(0.06, 0.4) +
  # adding the median separately as a point
  stat_summary(fun=mean, geom="point", shape=16) +
  labs(x="Tone types", y = "Syllable duration")
  #scale_y_log10()+

ggsave("Plot of duration by tone types.png")

aggregate(dur_by_tone$syllable_duration, list(dur_by_tone$syllable_tone), FUN=mean)
```

Intensity:
```{r}
int_by_tone <- syllable %>%
  group_by(syllable_tone, syllable_id) %>%
  summarise(syllable_intensity=syllable_intensity[1]) %>%
  ungroup()

ggplot(int_by_tone, aes(x=syllable_tone, y=syllable_intensity, fill=syllable_tone)) +
  geom_boxplot() +
  #ylim(0.06, 0.4) +
  # adding the median separately as a point
  stat_summary(fun=mean, geom="point", shape=16) +
  labs(x="Tone types", y = "Intensity")
  #scale_y_log10()+

ggsave("Plot of intensity by tone types.png")

aggregate(int_by_tone$syllable_intensity, list(int_by_tone$syllable_tone), FUN=mean)
```

Then I want to see how the duration of initial and final syllables look like for each tone group
```{r}
dur_by_tone_group <- syllable %>%
  #add tone value columns
  mutate(first_tone = if_else(position_in_word == 'initial', syllable_tone, 'NA')) %>%
  mutate(second_tone = if_else(position_in_word == 'final', syllable_tone, 'NA')) %>%
  
  #add duration columns
  mutate(first = if_else(position_in_word == 'initial', syllable_duration, 0)) %>%
  mutate(second = if_else(position_in_word == 'final', syllable_duration, 0)) %>%
  

  #group by words
  group_by(word) %>%
  #summarize 1st and 2nd for each word, and calculate mean duration for initial & second syllable for each word
  #summarise_at(vars(first), list(name = mean)) %>%
  summarise(first_tone=first_tone[1], second_tone = second_tone[2], first = mean(first) * 2, second = mean(second) *2)  %>%
  ungroup()
  
#view(dur_by_tone_group)  

#add a tone group column for each word
dur_by_tone_group$tone_group <- paste(dur_by_tone_group$first_tone, dur_by_tone_group$second_tone)

#view(dur_by_tone_group)




#list(dur_by_tone_group$tone_group)
#unique(dur_by_tone_group[c("tone_group")])
```
For now, remove all words that have NA (don't know why lots of tones are not encoded) --- each removes 2~3% rows
```{r}
dur_by_tone_group <- dur_by_tone_group %>%
  filter(!first_tone %in% c('NA')) %>%
  filter(!second_tone %in% c('NA')) %>%
  drop_na() %>%
  filter(!tone_group %in% c('t5 t1', 't5 t2', 't5 t3', 't5 t4'))

#view(dur_by_tone_group)
```


```{r}
dur_by_tone_group <- dur_by_tone_group %>%
  #group by tone group
  group_by(tone_group)
  #ungroup()
table(dur_by_tone_group$tone_group)
#view(dur_by_tone_group)

#melt the data, create a new column encode first/second syll's duration, and a value column
dur_by_tone_group <- melt(dur_by_tone_group)
#view(dur_by_tone_group)


ggplot(dur_by_tone_group, aes(x = variable, y = value, fill=variable)) +
  facet_wrap(~tone_group) +
  geom_boxplot() +
  stat_summary(fun=mean, geom="point", shape=16, size = 1) +
  ylim(0.1,0.3) +
  labs(x="Syllable position", y = "Syllable duration")
ggsave("Plot of duration by tone group.png")

```

Then I want to see how the intensity of initial and final syllables look like for each tone group
```{r}
int_by_tone_group <- syllable %>%
  #add tone value columns
  mutate(first_tone = if_else(position_in_word == 'initial', syllable_tone, 'NA')) %>%
  mutate(second_tone = if_else(position_in_word == 'final', syllable_tone, 'NA')) %>%
  
  #add intensity columns
  mutate(first = if_else(position_in_word == 'initial', syllable_intensity, 0)) %>%
  mutate(second = if_else(position_in_word == 'final', syllable_intensity, 0)) %>%

  #group by words
  group_by(word) %>%
  #summarize 1st and 2nd for each word, and calculate mean intensity for initial & second syllable for each word
  summarise(first_tone=first_tone[1], second_tone = second_tone[2], first = mean(first) * 2, second = mean(second) * 2)  %>%
  ungroup()
  
#add a tone group column for each word
int_by_tone_group$tone_group <- paste(int_by_tone_group$first_tone, int_by_tone_group$second_tone)

#view(int_by_tone_group)




#list(dur_by_tone_group$tone_group)
#unique(dur_by_tone_group[c("tone_group")])
```
For now, remove all words that have NA (don't know why lots of tones are not encoded) --- each removes 2~3% rows
(i.e. same words as in the previous duration measurements )
```{r}
int_by_tone_group <- int_by_tone_group %>%
  filter(!first_tone %in% c('NA')) %>%
  filter(!second_tone %in% c('NA')) %>%
  drop_na() %>%
  filter(!tone_group %in% c('t5 t1', 't5 t2', 't5 t3', 't5 t4'))
#filter tone group that are 51, 52, 53, 54

#view(int_by_tone_group)
```

Now it's time to encode first/second syllable duration for each word
```{r}
int_by_tone_group <- int_by_tone_group %>%
  #group by tone group
  group_by(tone_group)
  #ungroup()


#melt the data, create a new column encode first/second syll's duration, and a value column
int_by_tone_group <- melt(int_by_tone_group)
#view(int_by_tone_group)


ggplot(int_by_tone_group, aes(x = variable, y = value, fill=variable)) +
  facet_wrap(~tone_group) +
  geom_boxplot() +
  stat_summary(fun=mean, geom="point", shape=16, size = 1) +
  ylim(60,80) +
  labs(x="Syllable position", y = "Intensity")

ggsave("Plot of intensity by tone group.png")

```

## look at t3t5 and t2t5 tone groups: what lexical items are included
```{r}
int_by_tone_group <- int_by_tone_group %>%
  filter(tone_group %in% c('t3 t5')) %>%
  summarize(word, tone_group)

#view(int_by_tone_group)
```